---
title: "PKM-AI"
author: "Wilbert"
date: "2024-02-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load library
```{r}
library(coda)
library(rjags)
library(knitr)
library(purrr)
library(dplyr)
library(tidyr)
library(ggplot2)
```

# set seed
```{r}
set.seed(123)
```

# Load Dataset
```{r}
df <- read.csv("./Global_Education.csv")
show(df)
```

# Prepare Dataset
## Set 0 as missing value

```{r}
df[sapply(df, function(x) as.character(x) %in% c("0"))] <- NA
```

## Change Missing Value to median
```{r}
# Calculate medians for all numeric columns
numeric_columns <- sapply(df, is.numeric)
column_medians <- sapply(df[, numeric_columns], median, na.rm = TRUE)

# Replace NA values in numeric columns with their respective medians
for (col in names(column_medians)) {
    df[, col][is.na(df[, col])] <- column_medians[col]
}
```


## Drop useless variable
```{r}
df$Countries.and.areas <- NULL
```

## scale dataset
```{r}
df$Unemployment_Rate <- df$Unemployment_Rate / 100
head(df)
```
```{r}
for (x in colnames(df)) {
  cat(x, sep = "", ", ")
}
```


## Split features adn target
```{r}
Y <- as.matrix(df$Unemployment_Rate)
X <- cbind(df$OOSR_Pre0Primary_Age_Male, df$OOSR_Pre0Primary_Age_Female, df$OOSR_Primary_Age_Male, df$OOSR_Primary_Age_Female, df$OOSR_Lower_Secondary_Age_Male, df$OOSR_Lower_Secondary_Age_Female, df$OOSR_Upper_Secondary_Age_Male, df$OOSR_Upper_Secondary_Age_Female, df$Completion_Rate_Primary_Male, df$Completion_Rate_Primary_Female, df$Completion_Rate_Lower_Secondary_Male, df$Completion_Rate_Lower_Secondary_Female, df$Completion_Rate_Upper_Secondary_Male, df$Completion_Rate_Upper_Secondary_Female, df$Grade_2_3_Proficiency_Reading, df$Grade_2_3_Proficiency_Math, df$Primary_End_Proficiency_Reading, df$Primary_End_Proficiency_Math, df$Lower_Secondary_End_Proficiency_Reading, df$Lower_Secondary_End_Proficiency_Math, df$Youth_15_24_Literacy_Rate_Male, df$Youth_15_24_Literacy_Rate_Female, df$Gross_Primary_Education_Enrollment, df$Gross_Tertiary_Education_Enrollment)
names <- c("Intercepst","OOSR_Pre0Primary_Age_Male", "OOSR_Pre0Primary_Age_Female", "OOSR_Primary_Age_Male", "OOSR_Primary_Age_Female", "OOSR_Lower_Secondary_Age_Male", "OOSR_Lower_Secondary_Age_Female", "OOSR_Upper_Secondary_Age_Male", "OOSR_Upper_Secondary_Age_Female", "Completion_Rate_Primary_Male", "Completion_Rate_Primary_Female", "Completion_Rate_Lower_Secondary_Male", "Completion_Rate_Lower_Secondary_Female", "Completion_Rate_Upper_Secondary_Male", "Completion_Rate_Upper_Secondary_Female", "Grade_2_3_Proficiency_Reading", "Grade_2_3_Proficiency_Math", "Primary_End_Proficiency_Reading", "Primary_End_Proficiency_Math", "Lower_Secondary_End_Proficiency_Reading", "Lower_Secondary_End_Proficiency_Math", "Youth_15_24_Literacy_Rate_Male", "Youth_15_24_Literacy_Rate_Female",  "Gross_Primary_Education_Enrollment", "Gross_Tertiary_Education_Enrollment")
```

## Remove Missing Values
```{r}
junk <- is.na(rowSums(X))
Y <- Y[!junk]
X <- X[!junk,]
```

## Set X as Matrix
```{r}
X <- as.matrix(scale(X))
```

## Set Final Params
```{r}
X <- cbind(1,X) # add the intercept
colnames(X) <- names
n <- length(Y)
p <- ncol(X)
data <- list(Y=Y,X=X,n=n,p=p)
params <- c("beta","r")
```

# Model 1 
```{r}
model_string <- textConnection("model{
for(i in 1:n){
  Y[i] ~ dbeta(r*mu[i],r*(1-mu[i]))
  logit(mu[i]) <- inprod(X[i,],beta[])
  like[i] <- dbeta(Y[i], r*mu[i], r*(1-mu[i]))
}
for(j in 1:p){
  beta[j] ~ dnorm(0,0.01)
  } 
  r ~ dgamma(0.1,0.1)
}")
```

## Fit Model
```{r}
model <- jags.model(model_string,data = data, n.chains=2,quiet=TRUE)
```

## Update Model
```{r}
update(model, 10000)
```

## Generate Samples
```{r}
samples <- coda.samples(model, variable.names = params, n.iter = 75000, thin = 5)
```

## Evaluate Model
### Sample Plot
```{r}
par(mar=c(1,1,1,1))
plot(samples)
```

### Check Convergence
```{r}
gelman.diag(samples)
```

### Compile results
```{r}
ESS <- effectiveSize(samples)

ESS
```

```{r}
sum <- summary(samples)
rownames(sum$statistics) <- c(names,"r")
rownames(sum$quantiles) <- c(names,"r")
sum$statistics <- round(sum$statistics,3)
sum$quantiles <- round(sum$quantiles,3)
sum

```

### Autocorr Plot
```{r}
autocorr.plot(samples)
```


### DIC-WAIC Computation
```{r}
#DIC
#dic <- dic.samples(model,n.iter=10000)
```


```{r}
#dic1
```










